NIHAO VII: Predictions for the galactic baryon budget in dwarf to Milky Way mass haloes

### 2016-11-18 ms_nihao_budget.tex
first version according to the third referee report

\begin{abstract}
  We use the NIHAO galaxy formation simulations to make predictions
  for the baryonic budget  in present day galaxies ranging from dwarf
  ($\Mhalo\sim10^{10} \Msun$) to Milky Way ($\Mhalo\sim10^{12} \Msun$)
  masses.  The sample is made of 88 independent high resolution
  cosmological zoom-in simulations.  NIHAO galaxies reproduce key
  properties of observed galaxies, such as the stellar mass vs halo
  mass and cold gas vs stellar mass relations. Thus they make
  plausible predictions for the baryon budget.  We present the mass
  fractions of stars, cold gas ($T<10^4$K), cool gas ($10^4 < T <
  10^5$K), warm-hot gas ($10^5 < T < 10^7$K), and hot gas (T$> 10^7$K)
  , inside the virial radius, $R_{200}$.  Compared to the predicted
  baryon mass, using the dark halo mass and the universal baryon
  fraction, $f_{\rm b}\equiv \Omega_{\rm b}/\Omega_{\rm m}=0.15$, we
  find that all of our haloes are missing baryons. The missing mass
  has been relocated past 2 virial radii, and cool gas dominates
    the corona at low mass (M$_{200} \lta 3 \times 10^{11} \Msun$)
    while the  warm-hot gas dominates at high mass (M$_{200}
    \gta 3 \times 10^{11} \Msun$).  Haloes of mass $\Mhalo\sim
  10^{10}\Msun$ are missing $\sim 90\%$ of their baryons.  More
  massive haloes ($\Mhalo\sim 10^{12}\Msun$) retain a higher fraction
  of their baryons, with $\sim 30\%$ missing, consistent with recent
  observational estimates.  Moreover, these more massive haloes
    reproduce the observed  fraction of cold, warm-hot and hot gas.
    The fraction of cool gas we predict (0.11) is significantly lower
    than the observation from COS-HALOs (0.3-0.47), but agrees with
    the alternative analysis of \citet{Stern16}.
\end{abstract}

@@

{\bf As a part of the IGM, the circum galactic medium (CGM) is always 
treated as }an major potential
reservoir of the missing baryons. \hi{} and metal absorption lines are
expected  to signpost such diffuse baryonic content. \hi{} is mainly
from gas with temperature T $\sim 10^4$ K so that it is able to detect
cold gas in the CGM.  Meanwhile, theoretical work has predicted that a
substantial  portion of the CGM is in the warm or hot phase with
temperature T $> 10^{4.5}$ K.  Gas enters this phase through
photoionization, accretion shocks or shocks caused by galactic winds
\citep{Voort12}. Such dilute halo gas is at T$\sim 10^{4.5-7}$ K, so
the detection is dominated by metal lines, e.g. \ovi{}.  Recent
advances in the detection of gas in the CGM have come  from the COS
survey \citep{Tumlinson11, Tumlinson13, Thom12,  Werk12, Werk13}.  On
the scale of Milky Way mass haloes $\Mhalo \sim 10^{12}\Msun$ a
significant amount of warm ($10^4 < T < 10^7$K) gas has been detected
\citep{Werk14}, accounting for 33-88\% of the baryon budget. In the
future such observations will be extended to a wider range of halo
masses.  A number of large volume cosmological simulations
\citep{Ford13, Ford15, Suresh15b, Oppenheimer16} and zoom-in
cosmological simulations \citep{Stinson12, Hummels13, Shull14} have
given  predictions for the \hi{} and \ovi{} absorption
line. \citet{Gutcke16} compared the column density profile of \ovi{}
and \hi{} in the CGM of galaxies from the NIHAO \citep{Wang15}
cosmological hydrodynamical simulation suite with observations,
studied the covering fraction of dense \hi{}, looked at the shape of
the CGM and its chemical composition. The simulations
  reproduce  the observational covering fraction and column density
  profile of cool \hi{} well, and recover the observed trends of
  \ovi{} column density with luminosity and impact parameter. In
  common with other simulations, however, the column density of \ovi{}
  is lower and the extent of optically thick \hi{} is smaller than
  observed.

@@

@@

@@

In summary, while the stellar mass fractions are dependent on the
sub-grid models, the baryon fractions (for haloes in the mass range
$10^{10}\lta M_{200} \lta 10^{12}\Msun$) appear insensitive to the
details of the simulation code. In particular all three codes predict a
greater fraction of missing baryons in lower mass haloes. This
suggests that the potential well of the halo regulates the baryon
fraction more than the strength of the feedback.

@@

In Fig.~\ref{fig:corona}, we present the radial distribution of
gas in different phases at $z=0$, normalized to the total baryonic
mass profile, such that, in the region far away from the central galaxy
where the stars are rare, for a given halo the four phases add up to
unity, each line is colour coded by the virial mass.  
All simulations share a common attribute.  The cold gas (T$<10^4$K)
is mostly located  near the center  ($R<0.2R_{200}$)
where most stars in galaxies form.  In contrast, the cool ($10^4$K$<$T$<10^5$K)
and  warm-hot ($10^5$K$<$T$<10^7$K) gas are
located at large distances with roughly constant fractions up to 2
times $R_{200}$.  The hot gas (T$>10^7$K) is a minority component
for all galaxies in the NIHAO sample, with the maximum hot gas
fraction at any radius being  less than 5\%.

@@

The COS-HALOs survey is filling in details about the CGM at low
redshift \citep{Peeples14, Tumlinson11, Tumlinson13, Werk12, Werk13,
  Werk14}.  For the CGM of low-redshift $\sim L^*$ galaxies
($\Mstar\sim 10^{10.5}\Msun$), \citet{Tumlinson13} and
\citet{Peeples14} constrain the mass of  the warm-hot CGM ($T \sim
10^{5-7}$K), \citet{Werk14} provides a strict lower limit to the mass
of cool material ($T \sim 10^{4-5}$K) in the CGM of these galaxies.
In a study using X-rays, \citet{Anderson13} place a constraints on the
mass of  hot gas ($T > 10^7$K) residing in the extended hot halos.
These observational constraints are shown in Fig.~\ref{fig:comparison}
with the same colour scheme as in Fig.11 of \citet{Werk14}.  The
stellar mass fractions are based on halo abundance matching as
described in \citet{Kravtsov14}, while the cold disk gas mass comes
from \citet{Dutton11}.
{\bf The mean value of missing fraction is the residual fraction
of the sum of mean values of all components fractions in unity
and the lower and upper limits are calculated in the same manner
by the upper and lower limits of the all fractions.}

@@

The blue points and error bars show the mean values and standard
deviation of the mass fraction of stars and different components of
gas in our most massive galaxies ($3.49 \times 10^{11} \Msun$ $<$
M$_{200}$ $<$  $3.53 \times 10^{12} \Msun$).  The gas is assigned to a
range of temperature bins:  cold gas (T $< 10^4$ K), cool gas ($10^4$
K $<$ T $< 10^5$ K),  warm gas ($10^5$ K $<$ T $< 10^7$ K) and hot gas
(T $> 10^7$ K).  The observations and the simulations match well in
every phase except the cool CGM gas, where the observations from
\citet{Werk14} find $3\times$ the mass that NIHAO simulations predict.

@@

\item Cold gas ($T<10^4$K) is mostly restricted to be within 0.2
  virial radii (Fig.~\ref{fig:corona}). Cool gas ($10^4 < T < 10^5$K)
  dominates the baryonic mass outside the virial radius, {\bf as well
  as outside the 20\% $\Rvir$}, at low masses
  ($M_{200}\lta 3\times 10^{11} \Msun$) while the warm-hot gas ($10^5
  < T < 10^7$K) dominates at high masses ($M_{200}\gta 3\times 10^{11}
  \Msun$) (Figs.~\ref{fig:corona} \& \ref{fig:inout}).

@@

%% TABLE 1
\begin{table*}
  \caption{The baryonic budget parameters for NIHAO galaxies
    in different halo  mass bins. We refer to gas in the temperature range  T
  $<$ $10^4$ K as cold; $10^4$ K $\leqslant$ T $<$ $10^5$ K as cool;
  $10^5$ K $\leqslant$ T $<$ $10^7$ K as warm;  and T $\geqslant$
  $10^7$ K as hot.}
\begin{center}
\begin{tabular}{ccccc}
\hline
\input{budgettable4.txt}
\hline
\end{tabular}
\label{tab:comparison}
\end{center}
\end{table*}

@@

>>>>

\begin{abstract}
  We use the NIHAO galaxy formation simulations to make predictions
  for the baryonic budget  in present day galaxies ranging from dwarf
  ($\Mhalo\sim10^{10} \Msun$) to Milky Way ($\Mhalo\sim10^{12} \Msun$)
  masses.  The sample is made of 88 independent high resolution
  cosmological zoom-in simulations.  NIHAO galaxies reproduce key
  properties of observed galaxies, such as the stellar mass vs halo
  mass and cold gas vs stellar mass relations. Thus they make
  plausible predictions for the baryon budget.  We present the mass
  fractions of stars, cold gas ($T<10^4$K), cool gas ($10^4 < T <
  10^5$K), warm-hot gas ($10^5 < T < 5\times10^6$K), and hot gas (T$> 5\times10^6$K)
  , inside the virial radius, $R_{200}$.  Compared to the predicted
  baryon mass, using the dark halo mass and the universal baryon
  fraction, $f_{\rm b}\equiv \Omega_{\rm b}/\Omega_{\rm m}=0.15$, we
  find that all of our haloes are missing baryons. The missing mass
  has been relocated past 2 virial radii, and cool gas dominates
    the corona at low mass (M$_{200} \lta 3 \times 10^{11} \Msun$)
    while the  warm-hot gas dominates at high mass (M$_{200}
    \gta 3 \times 10^{11} \Msun$).  Haloes of mass $\Mhalo\sim
  10^{10}\Msun$ are missing $\sim 90\%$ of their baryons.  More
  massive haloes ($\Mhalo\sim 10^{12}\Msun$) retain a higher fraction
  of their baryons, with $\sim 30\%$ missing, consistent with recent
  observational estimates.  Moreover, these more massive haloes
    reproduce the observed  fraction of cold, warm-hot and hot gas.
    The fraction of cool gas we predict (0.11) is significantly lower
    than the observation from COS-HALOs (0.3-0.47), but agrees with
    the alternative analysis of \citet{Stern16}.
\end{abstract}

@@

{\bf As a part of the IGM, the circum galactic medium (CGM) is always 
treated as }an major potential
reservoir of the missing baryons. \hi{} and metal absorption lines are
expected  to signpost such diffuse baryonic content. \hi{} is mainly
from gas with temperature T $\sim 10^4$ K so that it is able to detect
cold gas in the CGM.  Meanwhile, theoretical work has predicted that a
substantial  portion of the CGM is in the warm or hot phase with
temperature T $> 10^{4.5}$ K.  Gas enters this phase through
photoionization, accretion shocks or shocks caused by galactic winds
\citep{Voort12}. Such dilute halo gas is at T$\sim 10^{4.5-7}$ K, so
the detection is dominated by metal lines, e.g. \ovi{}.  Recent
advances in the detection of gas in the CGM have come  from the COS
survey \citep{Tumlinson11, Tumlinson13, Thom12,  Werk12, Werk13}.  On
the scale of Milky Way mass haloes $\Mhalo \sim 10^{12}\Msun$ a
significant amount of warm ($10^4 < T <5\times 10^6$K) gas has been detected
\citep{Werk14}, accounting for 33-88\% of the baryon budget. In the
future such observations will be extended to a wider range of halo
masses.  A number of large volume cosmological simulations
\citep{Ford13, Ford15, Suresh15b, Oppenheimer16} and zoom-in
cosmological simulations \citep{Stinson12, Hummels13, Shull14} have
given  predictions for the \hi{} and \ovi{} absorption
line. \citet{Gutcke16} compared the column density profile of \ovi{}
and \hi{} in the CGM of galaxies from the NIHAO \citep{Wang15}
cosmological hydrodynamical simulation suite with observations,
studied the covering fraction of dense \hi{}, looked at the shape of
the CGM and its chemical composition. The simulations
  reproduce  the observational covering fraction and column density
  profile of cool \hi{} well, and recover the observed trends of
  \ovi{} column density with luminosity and impact parameter. In
  common with other simulations, however, the column density of \ovi{}
  is lower and the extent of optically thick \hi{} is smaller than
  observed.

@@

%% FIGURE 2
\begin{figure*}
\centerline{
  \psfig{figure=budget_res.eps,width=0.85\textwidth}
}
\caption{{\bf Linear fractional baryon content of NIHAO
simulations as a function of halo mass color coded by
number of dark matter particles per halo.}}
\label{fig:budget_res}
\end{figure*}

@@

Fig.~\ref{fig:budget_res} shows the budget fraction of cold, 
cool and warm-hot gas fraction. Since one can imagine that 
with increased resolution, higher densities can be achieved in the
CGM, potential leading to more cooling and larger
cold/cool gas fraction, all the points are color coded by 
the total number of dark matter particles within the haloes.
In the plot, the number of dark matter
partilce per halo spans 2 magnitudes and has no correlation 
with cold and cool gas fraction.
Therefore, the baryonic fractions predicted by NIHAO simulations
are convergent.

@@

In summary, while the stellar mass fractions are dependent on the
sub-grid models, the baryon fractions (for haloes in the mass range
$10^{10}\lta M_{200} \lta 10^{12}\Msun$) appear insensitive to the
details of the simulation code. In particular all three codes predict a
greater fraction of missing baryons in lower mass haloes. This
suggests that the potential well of the halo {\bf correlates the baryon
fraction more closely} than the strength of the feedback.

@@

In Fig.~\ref{fig:corona}, we present the radial distribution of
gas in different phases at $z=0$, normalized to the total baryonic
mass profile, such that, in the region far away from the central galaxy
where the stars are rare, for a given halo the four phases add up to
unity, each line is colour coded by the virial mass.  
All simulations share a common attribute.  The cold gas (T$<10^4$K)
is mostly located  near the center  ($R<0.2R_{200}$)
where most stars in galaxies form.  In contrast, the cool ($10^4$K$<$T$<10^5$K)
and  warm-hot ($10^5$K$<$T$<5\times10^6$K) gas are
located at large distances with roughly constant fractions up to 2
times $R_{200}$.  The hot gas (T$>5\times10^6$K) is a minority component
for all galaxies in the NIHAO sample, with the maximum hot gas
fraction at any radius being  less than 5\%.

@@

The COS-HALOs survey is filling in details about the CGM at low
redshift \citep{Peeples14, Tumlinson11, Tumlinson13, Werk12, Werk13,
  Werk14}.  For the CGM of low-redshift $\sim L^*$ galaxies
($\Mstar\sim 10^{10.5}\Msun$), \citet{Tumlinson13} and
\citet{Peeples14} constrain the mass of  the warm-hot CGM ($T \sim
10^{5-6.7}$K), \citet{Werk14} provides a strict lower limit to the mass
of cool material ($T \sim 10^{4-5}$K) in the CGM of these galaxies.
In a study using X-rays, \citet{Anderson13} place a constraints on the
mass of  hot gas ($T > 5\times10^6$K) residing in the extended hot halos.
These observational constraints are shown in Fig.~\ref{fig:comparison}
with the same colour scheme as in Fig.11 of \citet{Werk14}.  The
stellar mass fractions are based on halo abundance matching as
described in \citet{Kravtsov14}, while the cold disk gas mass comes
from \citet{Dutton11}.
{\bf In addition, since \citet{Crain10} and \citet{Voort13} show that 
soft X-ray emission traces gas with temperatures a few times the 
virial temperature (T$_{\rm vir} = 10^6$ K for M$_{200} \sim 
10^{12} \Msun$), we also estimate the hot gas fraction with lower
boundaries ($3\times 10^6$ K and $10^6$ K) between the hot gas and 
warm-hot gas, the fractions can increase to 0.00511 ($3\times 10^6$K) 
and 0.04168 ($10^6$ K).}
{\bf The mean value of missing fraction is the residual fraction
of the sum of mean values of all components fractions in unity
and the lower and upper limits are calculated in the same manner
by the upper and lower limits of the all fractions.}

@@

The blue points and error bars show the mean values and standard
deviation of the mass fraction of stars and different components of
gas in our most massive galaxies ($3.49 \times 10^{11} \Msun$ $<$
M$_{200}$ $<$  $3.53 \times 10^{12} \Msun$).  The gas is assigned to a
range of temperature bins:  cold gas (T $< 10^4$ K), cool gas ($10^4$
K $<$ T $< 10^5$ K),  warm gas ($10^5$ K $<$ T $<5\times 10^6$ K) and hot gas
(T $>5\times 10^6$ K).  The observations and the simulations match well in
every phase except the cool CGM gas, where the observations from
\citet{Werk14} find $3\times$ the mass that NIHAO simulations predict.

@@

\item Cold gas ($T<10^4$K) is mostly restricted to be within 0.2
  virial radii (Fig.~\ref{fig:corona}). Cool gas ($10^4 < T < 10^5$K)
  dominates the baryonic mass outside the virial radius, {\bf as well
  as outside the 20\% $\Rvir$}, at low masses
  ($M_{200}\lta 3\times 10^{11} \Msun$) while the warm-hot gas ($10^5
  < T <5\times 10^6$K) dominates at high masses ($M_{200}\gta 3\times 10^{11}
  \Msun$) (Figs.~\ref{fig:corona} \& \ref{fig:inout}).

@@

%% TABLE 1
\begin{table*}
  \caption{The baryonic budget parameters for NIHAO galaxies
    in different halo  mass bins. We refer to gas in the temperature range  T
  $<$ $10^4$ K as cold; $10^4$ K $\leqslant$ T $<$ $10^5$ K as cool;
  $10^5$ K $\leqslant$ T $<$ $5\times10^6$ K as warm;  and T $\geqslant$
  $5\times10^6$ K as hot.}
\begin{center}
\begin{tabular}{ccccc}
\hline
\input{budgettable4.txt}
\hline
\end{tabular}
\label{tab:comparison}
\end{center}
\end{table*}

@@

% Soft X-ray and ultraviolet metal-line emission from the gas around galaxies
\bibitem[van de Voort \& Schaye (2013)]{Voort13} van de Voort, F., Schaye, J.\ 2013, \mnras, 430, 2688
